---
title: "2007 Breeder Predictions Report"
author: "A-maize-ing Brains"
date: "2026-9-27"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(lme4)
library(lmerTest)
library(magrittr)
library(stringr)
library(caret)
library(scales)
library(ggpointdensity)
library(GGally)
library(patchwork)
knitr::opts_chunk$set(dev = 'png', dpi = 150, echo = TRUE)
```

# Trait Accuracy Report

This is a document that reports the overall accuracy of each model by presenting a heat map showing the quartile prediction accuracy. The models are trained on data from 2000-2006 and then tested on 2007 data. The train test split of predicting into a new year should give the most accurate representation of how our models will predict into a new year. In the heatmaps, the quartiles are listed from 1 to 4 with 4 being the lowest yielding genotypes and 1 being the highest yielding genotypes. Each square give the proportion of genotypes that fit into each quartile pair, with each column and row adding to 1. Additionally, accuracy statistics are split into 3 types:

1. Global Accuracy: These are calculated by simply including all predictions and then calculating rank accuracy
2. Genotypic Accuracy (G): These are calculated by averaging predictions across locations for each genotype and calculating rank accuracy
3. Within Environment Accuracy (GxE): These are calculated by calculating the rank accuracy within each location and then creating a weighted average across all locations based on the number of entries within each



```{r, echo=FALSE}
# Define Directories
datadir <- ""

modellist = c("GBLUP","XGB_BLUP","XGB_Env")
knitr::kable(modellist, caption = "Models")

traitnamelist = c("YLD_BE", "TWT","PHT")
bluplist <- c("_BLUP","_BLUP","")
knitr::kable(traitnamelist, caption = "Trait Predicted")



```


# Accuracy Statistics

## Yield

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE, out.width = "50%"}


i=1
t=1
d=1

 
for(w in 1:3){
 
  trait = traitnamelist[w]

  
  
  fulldf <- data.frame("Pedigree" = as.character())
  statstable <- data.frame("STAT" = c("Bot25","25p50p","50p75p","Top25","bot25inbot50","Top25intop50","top10","top10in25","rmse","corr"))
  
  
  ## Calculate Global Statistics First
  
  cat("\n\n\\pagebreak\n") 

  cat(paste(trait,"_Global Accuracy"))
  
  cat("\n")
  
  
  
for(i in 1:length(modellist)){

#Establish Model Type or Your Name
ModelID=modellist[i]

  
 scdf <- read.csv(here(datadir,paste0(modellist[i],"_Pred"),paste0("2007_",trait,bluplist[i],".csv")))
 colnames(scdf) <- c("LINE_UNIQUE_ID","Actual","Prediction")
 
 
 #Heatmap
 
  CombinedResults <- scdf
   CombinedResults <- CombinedResults %>% arrange(desc(Prediction)) %>% mutate(PredictionRank = row_number()/nrow(CombinedResults))%>% 
     arrange(desc(Actual)) %>% mutate(TraitRank = row_number()/nrow(CombinedResults))
  
#Create Percentile Accuracy Matrix
  
entry_n=count(CombinedResults)

fourfour=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.75) %>% 
  count())/entry_n/.25
fourthree=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
fourtwo=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
fourone=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank<.25) %>% 
  count())/entry_n/.25

threefour=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.75) %>% 
  count())/entry_n/.25
threethree=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
threetwo=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
threeone=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank<.25) %>% 
  count())/entry_n/.25

twofour=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.75) %>% 
  count())/entry_n/.25
twothree=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
twotwo=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
twoone=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank<.25) %>% 
  count())/entry_n/.25

onefour=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.75) %>% 
  count())/entry_n/.25
onethree=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
onetwo=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
oneone=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank<.25) %>% 
  count())/entry_n/.25

comparison_matrix=tibble(Prediction_Quartile=paste0("Q",c(rep(1:4,each=4))),True_Quartile=as.factor(paste0("Q",c(rep(1:4,times=4)))),
                          Value=as.numeric(c(oneone,onetwo,onethree,onefour,twoone,twotwo,twothree,twofour,threeone,threetwo,threethree,threefour,
                                  fourone,fourtwo,fourthree,fourfour)))


heatmap = ggplot(comparison_matrix, aes(Prediction_Quartile, True_Quartile)) +
  geom_tile(aes(fill = Value)) +
  geom_text(aes(label = round(Value, 4))) +
  scale_fill_gradient(low = "red", high = "green",
                      limits=c(0,0.6),oob=squish)+
  scale_y_discrete(limits = c("Q4","Q3","Q2","Q1"))+
  ggtitle(paste0("Accuracy_Matrix_2007_",ModelID,"_",trait))
print(heatmap)
 
 
# Calculate Statistics
df <- scdf

df <- df %>% arrange(desc(Prediction)) %>% mutate(PredictionRank = row_number()/nrow(df))%>% 
     arrange(desc(Actual)) %>% mutate(ActualRank = row_number()/nrow(df))

  Q4 <- df %>% filter(PredictionRank < 0.25 & ActualRank < 0.25)
Q4 <- 4*nrow(Q4)/nrow(df)

Q450 <- df %>% filter(PredictionRank < 0.50 & ActualRank < 0.25)
Q450 <- 4*nrow(Q450)/nrow(df)

Q1 <- df %>% filter(PredictionRank > 0.75 & ActualRank > 0.75)
Q1 <- 4*nrow(Q1)/nrow(df)

Q150 <- df %>% filter(PredictionRank > 0.50 & ActualRank > 0.75)
Q150 <- 4*nrow(Q150)/nrow(df)

Q2 <- df %>% filter(between(PredictionRank,0.50,0.75) & between(ActualRank,0.50,0.75))
Q2 <- 4*nrow(Q2)/nrow(df)

Q3 <- df %>% filter(between(PredictionRank,0.25,0.5) & between(ActualRank,0.25,0.5))
Q3 <- 4*nrow(Q3)/nrow(df)

top10 <- df %>% filter(PredictionRank < 0.1 & ActualRank < 0.1)
top10 <- 10*nrow(top10)/nrow(df)

top1025 <- df %>% filter(PredictionRank < 0.25 & ActualRank < 0.1)
top1025 <- 10*nrow(top1025)/nrow(df)

rmse <- df %>% mutate(resid = (PredictionRank - ActualRank)^2) %>% pull(resid) %>% mean() %>% sqrt

corr <- cor(df$PredictionRank,df$ActualRank)

globalstat <- c(Q1,Q2,Q3,Q4,Q150,Q450,top10,top1025,rmse,corr) %>% round(4)
 
 statstable <- cbind(statstable,globalstat)
 
}
  
  colnames(statstable) <- c("STAT",modellist)
  print(knitr::kable(statstable, caption = "Model Correlation Statistics"))
  
 
   ## Calculate Genetic Statistics
  
  cat("\n\n\\pagebreak\n") 

  cat(paste(trait," Genetic Accuracy"))
  
  cat("\n")
  
  statstable <- data.frame("STAT" = c("Bot25","25p50p","50p75p","Top25","bot25inbot50","Top25intop50","top10","top10in25","rmse","corr"))
  
for(i in 1:length(modellist)){

#Establish Model Type or Your Name
ModelID=modellist[i]

  
  
  scdf <- read.csv(here(datadir,paste0(modellist[i],"_Pred"),paste0("2007_",trait,bluplist[i],".csv")))
  colnames(scdf) <- c("LINE_UNIQUE_ID","Actual","Prediction")
  
  scdf <- scdf |> group_by(LINE_UNIQUE_ID) %>% summarize(Actual = mean(Actual), Prediction = mean(Prediction)) %>% ungroup()
 
 
 
 #Heatmap
 
  CombinedResults <- scdf
   CombinedResults <- CombinedResults %>% arrange(desc(Prediction)) %>% mutate(PredictionRank = row_number()/nrow(CombinedResults))%>% 
     arrange(desc(Actual)) %>% mutate(TraitRank = row_number()/nrow(CombinedResults))
  
#Create Percentile Accuracy Matrix
  
entry_n=count(CombinedResults)

fourfour=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.75) %>% 
  count())/entry_n/.25
fourthree=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
fourtwo=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
fourone=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank<.25) %>% 
  count())/entry_n/.25

threefour=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.75) %>% 
  count())/entry_n/.25
threethree=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
threetwo=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
threeone=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank<.25) %>% 
  count())/entry_n/.25

twofour=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.75) %>% 
  count())/entry_n/.25
twothree=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
twotwo=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
twoone=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank<.25) %>% 
  count())/entry_n/.25

onefour=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.75) %>% 
  count())/entry_n/.25
onethree=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
onetwo=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
oneone=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank<.25) %>% 
  count())/entry_n/.25

comparison_matrix=tibble(Prediction_Quartile=paste0("Q",c(rep(1:4,each=4))),True_Quartile=as.factor(paste0("Q",c(rep(1:4,times=4)))),
                          Value=as.numeric(c(oneone,onetwo,onethree,onefour,twoone,twotwo,twothree,twofour,threeone,threetwo,threethree,threefour,
                                  fourone,fourtwo,fourthree,fourfour)))


heatmap = ggplot(comparison_matrix, aes(Prediction_Quartile, True_Quartile)) +
  geom_tile(aes(fill = Value)) +
  geom_text(aes(label = round(Value, 4))) +
  scale_fill_gradient(low = "red", high = "green",
                      limits=c(0,0.6),oob=squish)+
  scale_y_discrete(limits = c("Q4","Q3","Q2","Q1"))+
  ggtitle(paste0("Accuracy_Matrix_2007_",ModelID,"_",trait))
print(heatmap)
 
 
# Calculate Statistics
df <- scdf

df <- df %>% arrange(desc(Prediction)) %>% mutate(PredictionRank = row_number()/nrow(df))%>% 
     arrange(desc(Actual)) %>% mutate(ActualRank = row_number()/nrow(df))

  Q4 <- df %>% filter(PredictionRank < 0.25 & ActualRank < 0.25)
Q4 <- 4*nrow(Q4)/nrow(df)

Q450 <- df %>% filter(PredictionRank < 0.50 & ActualRank < 0.25)
Q450 <- 4*nrow(Q450)/nrow(df)

Q1 <- df %>% filter(PredictionRank > 0.75 & ActualRank > 0.75)
Q1 <- 4*nrow(Q1)/nrow(df)

Q150 <- df %>% filter(PredictionRank > 0.50 & ActualRank > 0.75)
Q150 <- 4*nrow(Q150)/nrow(df)

Q2 <- df %>% filter(between(PredictionRank,0.50,0.75) & between(ActualRank,0.50,0.75))
Q2 <- 4*nrow(Q2)/nrow(df)

Q3 <- df %>% filter(between(PredictionRank,0.25,0.5) & between(ActualRank,0.25,0.5))
Q3 <- 4*nrow(Q3)/nrow(df)

top10 <- df %>% filter(PredictionRank < 0.1 & ActualRank < 0.1)
top10 <- 10*nrow(top10)/nrow(df)

top1025 <- df %>% filter(PredictionRank < 0.25 & ActualRank < 0.1)
top1025 <- 10*nrow(top1025)/nrow(df)

rmse <- df %>% mutate(resid = (PredictionRank - ActualRank)^2) %>% pull(resid) %>% mean() %>% sqrt

corr <- cor(df$PredictionRank,df$ActualRank)

globalstat <- c(Q1,Q2,Q3,Q4,Q150,Q450,top10,top1025,rmse,corr) %>% round(4)
 
 statstable <- cbind(statstable,globalstat)
 
}
  
  colnames(statstable) <- c("STAT",modellist)
  print(knitr::kable(statstable, caption = "Model Correlation Statistics"))
  
  
  
   ## Calculate GxE Statistics
  
  cat("\n\n\\pagebreak\n") 

  cat(paste(trait," Within Location Accuracy"))
  
  cat("\n")
  
  statstable <- data.frame("STAT" = c("Bot25","25p50p","50p75p","Top25","bot25inbot50","Top25intop50","top10","top10in25","rmse","corr"))
  
  i=3
  
for(i in 3){

  locdf <- read.csv(paste0("Phenotype_Files/",trait,".csv")) |> filter(YEAR == 2007) |> 
    select("LINE_UNIQUE_ID","LOC",trait)
  
  colnames(locdf) <- c("LINE_UNIQUE_ID","LOC","Actual")
  
#Establish Model Type or Your Name
ModelID=modellist[i]

  
 scdf <- read.csv(here(datadir,paste0(modellist[i],"_Pred"),paste0("2007_",trait,bluplist[i],".csv")))
  colnames(scdf) <- c("LINE_UNIQUE_ID","Actual","Prediction")
  
  scdf <- inner_join(scdf,locdf, by = join_by("LINE_UNIQUE_ID","Actual"))
  
 locidlist <- scdf %>% pull(LOC) %>% unique()
 
 tempdf <- data.frame("Value" = c(rep(0,times=16)))
 
 for(l in 1:length(locidlist)){
 
 #Heatmap
 
   locid <- locidlist[l]
   
  CombinedResults <- scdf %>% filter(LOC == locid)
   CombinedResults <- CombinedResults %>% arrange(desc(Prediction)) %>% mutate(PredictionRank = row_number()/nrow(CombinedResults))%>% 
     arrange(desc(Actual)) %>% mutate(TraitRank = row_number()/nrow(CombinedResults))
  
#Create Percentile Accuracy Matrix
  
entry_n=count(CombinedResults)

fourfour=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.75) %>% 
  count())/entry_n/.25
fourthree=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
fourtwo=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
fourone=(CombinedResults %>% 
  filter(PredictionRank>.75&TraitRank<.25) %>% 
  count())/entry_n/.25

threefour=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.75) %>% 
  count())/entry_n/.25
threethree=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
threetwo=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
threeone=(CombinedResults %>% 
  filter(PredictionRank>.50&PredictionRank<.75&TraitRank<.25) %>% 
  count())/entry_n/.25

twofour=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.75) %>% 
  count())/entry_n/.25
twothree=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
twotwo=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
twoone=(CombinedResults %>% 
  filter(PredictionRank>.25&PredictionRank<.50&TraitRank<.25) %>% 
  count())/entry_n/.25

onefour=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.75) %>% 
  count())/entry_n/.25
onethree=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.50&TraitRank<.75) %>% 
  count())/entry_n/.25
onetwo=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank>.25&TraitRank<.50) %>% 
  count())/entry_n/.25
oneone=(CombinedResults %>% 
  filter(PredictionRank<.25&TraitRank<.25) %>% 
  count())/entry_n/.25


temp <- as.numeric(c(oneone,onetwo,onethree,onefour,twoone,twotwo,twothree,twofour,threeone,threetwo,threethree,threefour,
                                  fourone,fourtwo,fourthree,fourfour))

tempdf <- tempdf %>% mutate(Value = Value + nrow(CombinedResults)/nrow(scdf)*temp)

}

comparison_matrix=tibble(Prediction_Quartile=paste0("Q",c(rep(1:4,each=4))),True_Quartile=as.factor(paste0("Q",c(rep(1:4,times=4)))),
                          Value=tempdf[,1])


heatmap = ggplot(comparison_matrix, aes(Prediction_Quartile, True_Quartile)) +
  geom_tile(aes(fill = Value)) +
  geom_text(aes(label = round(Value, 4))) +
  scale_fill_gradient(low = "red", high = "green",
                      limits=c(0,0.6),oob=squish)+
  scale_y_discrete(limits = c("Q4","Q3","Q2","Q1"))+
  ggtitle(paste0("Accuracy_Matrix_2007_",ModelID,"_",trait))
print(heatmap)
 

tempstat <- data.frame("Value" = c(rep(0,times=10)))

for(l in 1:length(locidlist)){
 
# Calculate Statistics
  

  
  locid <- locidlist[l]
  
df <- scdf %>% filter(LOC == locid)

df <- df %>% arrange(desc(Prediction)) %>% mutate(PredictionRank = row_number()/nrow(df))%>% 
     arrange(desc(Actual)) %>% mutate(ActualRank = row_number()/nrow(df))

  Q4 <- df %>% filter(PredictionRank < 0.25 & ActualRank < 0.25)
Q4 <- 4*nrow(Q4)/nrow(df)

Q450 <- df %>% filter(PredictionRank < 0.50 & ActualRank < 0.25)
Q450 <- 4*nrow(Q450)/nrow(df)

Q1 <- df %>% filter(PredictionRank > 0.75 & ActualRank > 0.75)
Q1 <- 4*nrow(Q1)/nrow(df)

Q150 <- df %>% filter(PredictionRank > 0.50 & ActualRank > 0.75)
Q150 <- 4*nrow(Q150)/nrow(df)

Q2 <- df %>% filter(between(PredictionRank,0.50,0.75) & between(ActualRank,0.50,0.75))
Q2 <- 4*nrow(Q2)/nrow(df)

Q3 <- df %>% filter(between(PredictionRank,0.25,0.5) & between(ActualRank,0.25,0.5))
Q3 <- 4*nrow(Q3)/nrow(df)

top10 <- df %>% filter(PredictionRank < 0.1 & ActualRank < 0.1)
top10 <- 10*nrow(top10)/nrow(df)

top1025 <- df %>% filter(PredictionRank < 0.25 & ActualRank < 0.1)
top1025 <- 10*nrow(top1025)/nrow(df)

rmse <- df %>% mutate(resid = (PredictionRank - ActualRank)^2) %>% pull(resid) %>% mean() %>% sqrt

corr <- cor(df$PredictionRank,df$ActualRank)

globalstat <- as.numeric(c(Q1,Q2,Q3,Q4,Q150,Q450,top10,top1025,rmse,corr)) %>% round(4)
 
tempstat <- tempstat %>% mutate(Value = Value + nrow(df)/nrow(scdf)*globalstat)

 
}

statstable <- cbind(statstable,tempstat[,1])
  
}
  
  colnames(statstable) <- c("STAT","XGB_Env")
  print(knitr::kable(statstable, caption = "Model Correlation Statistics"))
  
}
  
```


