---
title: "Bayer Yield Stability Environment Paramaterization"
author: "Chris Barron"
date: "2025-07-24"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(here)
library(dplyr)
library(tidyverse)
library(ggfortify)
library(ggplot2)
library(readxl)
library(openxlsx)
library(ggpubr)
library(rstatix)
library(lme4)
library(lmerTest)
library(car)

knitr::opts_chunk$set(echo = TRUE)

datadir <- "Data/Bayer_Yield_Rsheet.csv"
weatherdir <- "Data/Bayer_Weather_Rsheet.csv"

read_excel_allsheets <- function(filename, tibble = FALSE) {
    sheets <- readxl::excel_sheets(filename)
    x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
    if(!tibble) x <- lapply(x, as.data.frame)
    names(x) <- sheets
    x
}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Correction

YLD - uncorrected yield
YLD_BE - Spatially adjusted grain yield

Data file also includes population and moisture, so it might be good to correct yield based on these too

Plant height and ear height traits are also available


### Experiments/Tasks

1) Parameterize Environments - Convert times-scale data into binned data based on GDD (assuming RM of 120) and also include static environmental variables.
  a. Determine if any of these variables are signficantly correlated to yield or if anova of environmental effects and yield has signficant explanations of vairance
  b. Assign environments as stressed.excess, stressed.lack, and optimum. I can play around with the specifics of these, perhaps with a scale of -3 to 3 where 0 is optimum, -3 is worst lack, and 3 is worst excess
  c. Develop an environment relationship/covariance matrix

2) After classifying environments as optimal or stressed, perform same regressions as for ACRE and GOTH where the trt effect is not explained by year. See if there is a singificant environment - year interaction

3) Based on environment classification, see if there are differences in the yield stability of different hybrids across environments


## Calcualte In Season and Cumulative GDD for each location - year

```{r}
# Load weather data
weatherdf <- read.csv(weatherdir)

# Load Yield Data, which inlcudes planting and harvest dates
yielddf <- read.csv(datadir)

# Extract Planting Dates
plantingdf <- yielddf %>% select(locyear,planting_date) %>% unique() %>% mutate(plantday = as.numeric(str_extract(planting_date, "(?<=/)\\d+"))) %>% 
  mutate(plantmonth = as.numeric(str_extract(planting_date, "\\d+(?=/)"))) %>% mutate(planting_date_numeric = plantmonth + 0.01*plantday)

#Convert dates in weather data to numeric so they can be filtered
weatherdf <- weatherdf %>% mutate(day = as.numeric(str_extract(date, "(?<=/)\\d+"))) %>% 
  mutate(month = as.numeric(str_extract(date, "\\d+(?=/)"))) %>% mutate(date_numeric = month + 0.01*day) %>% 
  mutate(locyear = paste0(loc_id, "_",planting_year))

locyearlist <- weatherdf %>% pull(locyear) %>% unique()

#Calculate cumulative GDD for each locyr within the growing season

allGDD <- data.frame(matrix(ncol = 4, nrow=0))
colnames(allGDD) <- c("locyear", "date_numeric", "inseasonGDD", "cumulativeinseasonGDD")

for(l in 1:length(locyearlist)){
  
  locyr <- locyearlist[l]
  
  locdate <- plantingdf %>% filter(locyear == locyr) %>% pull(planting_date_numeric)
  
  if(length(locdate) != 1){
    
    next
    
  }
  
  locweatherdf <- weatherdf %>% filter(locyear == locyr) %>% filter(date_numeric >= locdate) %>%
    mutate(GDD_airmax = if_else(max_temperature > 86, 86, max_temperature)) %>% mutate(inseasonGDD = (GDD_airmax + min_temperature)/2 - 50)
  
  locweatherdf <- locweatherdf %>% mutate(inseasonGDD = if_else(inseasonGDD < 0, 0, inseasonGDD)) %>% arrange(date_numeric)
  
  locweatherdf <- locweatherdf %>% mutate(cumulativeinseasonGDD = cumsum(inseasonGDD))
  
  GDDvalues <- locweatherdf %>% select(locyear, date_numeric, inseasonGDD, cumulativeinseasonGDD )
  
  allGDD <- rbind(allGDD, GDDvalues)
  
}

#Two loc - years have missing weather data

newweatherdf <- left_join(weatherdf, allGDD, by = join_by("locyear","date_numeric"))

write.csv(newweatherdf, here("Data","Bayer_weather_inseasonGDD.csv"))

```



## Bin Weather by Growth Stage based on Relative Maturity of 110
```{r}
# Load weather data
weatherdf <- read.csv(here("Data","Bayer_weather_inseasonGDD.csv"))

# Load Yield Data, which inlcudes planting and harvest dates
yielddf <- read.csv(datadir)


 RM = 110

GDDdf <- yielddf %>% select(locyear) %>% mutate(EmergeGDD = 115) %>% 
  mutate(EarlyVeg1GDD = 1.5*RM+115) %>% 
  mutate(EarlyVeg2GDD = 4*RM+115) %>% 
  mutate(MidVegGDD = 8*RM+115) %>% 
  mutate(LateVegGDD = 10*RM+115) %>% 
  mutate(SilkGDD = 11.456*RM+100.57) %>% 
  mutate(BlisterGDD = 14.4*RM+77.409) %>% 
  mutate(DoughGDD = 16.8*RM+54.259) %>% 
  mutate(DentGDD = 19.2*RM+31.109) %>% 
  mutate(BLGDD = 24.158*RM-15.191)


locyrlist <- GDDdf %>% unique()

# Remove rows with NA values, these represent rows that have weather data before or after harvest
weatherdf <- na.omit(weatherdf)

weathercols <- colnames(weatherdf[,4:44])

lablesvector <- c("Emergence","EarlyVeg1","EarlyVeg2","MidVeg","LateVeg","Silk","Blister","Dough",
                    "Dent","BL")

combinedbinnedweather <- data.frame(matrix(ncol = 442, nrow = 0))

colnames(combinedbinnedweather) <- c(colnames(yielddf[,c(1:3,5,11,15,18,19,22:27,29:46)]),paste0(rep(weathercols,each=10),"_",rep(lablesvector,times=41)))

for(i in 1:nrow(locyrlist)){
  
  singleyield <- locyrlist[i,]
  
  locyr <- singleyield %>% pull(locyear)
  
  cutvector <- t(singleyield[1,2:11]) %>% as.vector()
  cutvector <- c(0,cutvector)

  
  redweatherdf <- weatherdf %>% filter(locyear == locyr) %>% 
    mutate(cumulativeinseasonGDD = as.numeric(cumulativeinseasonGDD))
 # Skip Entries where we do not have weather data for that loc year pair
  if(nrow(redweatherdf) == 0){
    
    next
  }
  
  redweatherdf <- redweatherdf %>% mutate(Maturity = cut(cumulativeinseasonGDD,
                                                         breaks=cutvector,
                                                         labels = lablesvector,
                                                         include.lowest=T))
  redweatherdf <- redweatherdf %>% group_by(Maturity) %>% summarize_at(all_of(weathercols),mean) %>% na.omit()
  
# Skip entries where not enough GDDs accumulated for all growth stages
  if(nrow(redweatherdf) != 10){
    
    next
    
  }
  
  weathervars <- redweatherdf %>% pivot_wider(names_from = Maturity, values_from = -c(Maturity))
  
  
  newentry <- cbind(singleyield,weathervars)
  
  combinedbinnedweather <- rbind(combinedbinnedweather,newentry)
}


entryandphenodf <- yielddf[,c(1:3,5,11,15,18,19,22:27,29:46)]

fullweatherentrydf <- inner_join(entryandphenodf,combinedbinnedweather,by=join_by("locyear"))


write.csv(fullweatherentrydf,here("Data","10GS_Binned_Weather_Data_Mean.csv"))

```

Old Weather file was binned based on 120 RM, new relative matrity file binned based on 110 RM



```{r}
weatherpcs <- prcomp(combinedbinnedweather[,c(12:81)])

autoplot(weatherpcs,
         data = combinedbinnedweather)
```

## Bin Weather by Growth Stage based on Relative Maturity of 110
```{r}
# Load weather data
weatherdf <- read.csv(here("Data","Bayer_weather_inseasonGDD.csv"))

# Load Yield Data, which inlcudes planting and harvest dates
yielddf <- read.csv(datadir)


 RM = 110

GDDdf <- yielddf %>% select(locyear) %>% mutate(EmergeGDD = 115) %>% 
  mutate(EarlyVeg1GDD = 1.5*RM+115) %>% 
  mutate(EarlyVeg2GDD = 4*RM+115) %>% 
  mutate(MidVegGDD = 8*RM+115) %>% 
  mutate(LateVegGDD = 10*RM+115) %>% 
  mutate(SilkGDD = 11.456*RM+100.57) %>% 
  mutate(BlisterGDD = 14.4*RM+77.409) %>% 
  mutate(DoughGDD = 16.8*RM+54.259) %>% 
  mutate(DentGDD = 19.2*RM+31.109) %>% 
  mutate(BLGDD = 24.158*RM-15.191)


locyrlist <- GDDdf %>% unique()

# Remove rows with NA values, these represent rows that have weather data before or after harvest
weatherdf <- na.omit(weatherdf)

weatherdf <- weatherdf %>% mutate(mean_temp = (max_temperature + min_temperature)/2) %>% 
  select(mean_temp,total_precipitation.inch.of.water,locyear,cumulativeinseasonGDD)

weathercols <- colnames(weatherdf[,1:2])

lablesvector <- c("Emergence","EarlyVeg1","EarlyVeg2","MidVeg","LateVeg","Silk","Blister","Dough",
                    "Dent","BL")

combinedbinnedweather <- data.frame(matrix(ncol = 31, nrow = 0))

colnames(combinedbinnedweather) <- c(colnames(yielddf[,c(1:3,5,11,22:27)]),paste0(rep(weathercols,each=10),"_",rep(lablesvector,times=2)))

for(i in 1:nrow(locyrlist)){
  
  singleyield <- locyrlist[i,]
  
  locyr <- singleyield %>% pull(locyear)
  
  cutvector <- t(singleyield[1,2:11]) %>% as.vector()
  cutvector <- c(0,cutvector)

  
  redweatherdf <- weatherdf %>% filter(locyear == locyr) %>% 
    mutate(cumulativeinseasonGDD = as.numeric(cumulativeinseasonGDD))
 # Skip Entries where we do not have weather data for that loc year pair
  if(nrow(redweatherdf) == 0){
    
    next
  }
  
  redweatherdf <- redweatherdf %>% mutate(Maturity = cut(cumulativeinseasonGDD,
                                                         breaks=cutvector,
                                                         labels = lablesvector,
                                                         include.lowest=T))
  redweatherdf <- redweatherdf %>% group_by(Maturity) %>% summarize_at(all_of(weathercols),mean) %>% na.omit()
  
# Skip entries where not enough GDDs accumulated for all growth stages
  if(nrow(redweatherdf) != 10){
    
    next
    
  }
  
  redweatherdf <- redweatherdf %>% mutate(deficit_temp = if_else(mean_temp <73.4,1,0)) %>% mutate(excess_temp = if_else(mean_temp>89.6,1,0)) %>% 
    mutate(deficit_precip = if_else(total_precipitation.inch.of.water<0.39,1,0)) %>%
    mutate(excess_precip = if_else(total_precipitation.inch.of.water > 0.78,1,0))
  
  weathervars <- redweatherdf %>% pivot_wider(names_from = Maturity, values_from = -c(Maturity))
  
  newentry <- cbind(singleyield,weathervars)
  
  combinedbinnedweather <- rbind(combinedbinnedweather,newentry)
}

combinedbinnedweather <- combinedbinnedweather %>% rowwise() %>%  mutate(total_excess = sum(c_across(starts_with("excess_")))) %>% 
  mutate(total_deficit = sum(c_across(starts_with("deficit_"))))

range(pull(combinedbinnedweather,total_deficit))

entryandphenodf <- yielddf[,c(1:3,5,11,22:27)]

fullweatherentrydf <- inner_join(entryandphenodf,combinedbinnedweather,by=join_by("locyear"))
```

# Comparison of Stress Covariates and Mean Yield

```{r}
yielddf <- read.csv(datadir)

genoyrvalues <- read_excel_allsheets(here("Data","Bayer2021ACRERsheet.xlsx"))
genoyrvalues <- genoyrvalues[[3]]
genoyrvalues <- genoyrvalues %>% select(Genotype,Year) %>% mutate(Year = Year - 1982) %>% unique()

yielddf <-  left_join(yielddf,genoyrvalues,by = "Genotype")

# Remove outliers

dfouts <- yielddf %>% mutate_if(is.numeric,scale) %>% mutate(across(is.numeric,~ifelse(abs(.)>4,NA,1)))

yielddf[22:27] <- yielddf[22:27]*dfouts[22:27]


locsumdf <- yielddf %>% select(locyear,YLD) %>% group_by(locyear) %>% summarize(MeanYLD = mean(YLD, na.rm=TRUE)) %>% ungroup()

stressdf <- fullweatherentrydf %>% select(locyear,total_deficit)

locsumdf <- inner_join(stressdf,locsumdf,by="locyear")

plot3 <- ggplot(locsumdf, aes(x=MeanYLDZ, y=total_deficit))+
  ylab("Total Stress by Deficit")+
  geom_point(size=2)
print(plot3)
```



# ANalysis of within year-loc stdev of yield

```{r}
df <- fullweatherentrydf[,c(1:3,7,83)]

genoyrvalues <- read_excel_allsheets(here("Data","Bayer2021ACRERsheet.xlsx"))
genoyrvalues <- genoyrvalues[[3]]
genoyrvalues <- genoyrvalues %>% select(Genotype,Year) %>% mutate(Year = Year - 1982) %>% unique()

df <-  left_join(df,genoyrvalues,by = "Genotype")

df <- df %>% mutate(Class = if_else(total_deficit < 13, "Optimal", if_else(total_deficit < 15, "Mid", "High")))

temp <- df %>% count(Class)
hist(df$total_deficit)

df <- df %>% na.omit() %>% group_by(Genotype, Year, Class) %>% summarize(Yld_Sdtev = var(YLD_BE)) %>% na.omit() %>% ungroup()

traits = colnames(df[,c(3:4)])

for (t in 2){

trait <- traits[t]

cat("\n\n\\pagebreak\n") 
  cat(paste(trait," With GS as Factor no SPATs"))
  cat("\n")

traitdf <- df %>%  select(c("Class","Genotype","Year",trait)) %>% na.omit() %>% rename("Trait" = trait) %>%
  mutate(Year = as.numeric(Year)) %>% mutate(Class = as.factor(Class))

#Visualize the Data

bxp2 <- ggboxplot(traitdf, x="Class", y="Trait",color = "Class",lwd = 1)+
  ylab(trait)
print(bxp2)

# Check Normality Assumption

normmodel <- lm(Trait~Year+factor(Class)+Year*factor(Class),data = traitdf)
normplot <-  ggqqplot(residuals(normmodel))+
   ggtitle(paste0(trait," QQ Plot"))

print(normplot)

shapiropval <- shapiro.test(normmodel$residuals)
shapiropval <- shapiropval$p.value %>% round(4)


#Constant variance


Homoscedasticity <- ncvTest(normmodel)
Homoscedasticityp <- Homoscedasticity$p %>% round(4)

#If p-value is not significant, variance is constant

#Anova test for variety and replication

anovatest <- anova_test(data = traitdf, Trait~Year+factor(Class)+Year*factor(Class))

anovatest <- as.tibble(anovatest)

# Determine if any of the effects are significant
anovanames <- anovatest %>% pull(Effect)
anovapvals <- anovatest %>% pull(p)

#Fit regression across treatments
modelall <- lm(Trait~ Year + factor(Class)+Year*factor(Class),traitdf)
summaryall <- summary(modelall)

predictions <- predict(modelall,traitdf)
residdf <- data.frame("residuals" = summaryall$residuals, "predictions" = predictions, "year" = traitdf$Year)
residplot <- ggplot(residdf, aes(x = predictions, y=residuals))+
  geom_point()

coefsall <- summaryall$coefficients

CoefHigh <- sum(coefsall[c(2),1]) %>% round(4)
CoefHighp <- if_else(min(coefsall[c(2),4]) <= 0.05, "*", "")

CoefMid <- sum(coefsall[c(2,3,5),1]) %>% round(4)
CoefMidp <- if_else(min(coefsall[c(2,3,5),4]) <= 0.05, "*", "")

CoefOptim <- sum(coefsall[c(2,4,6),1]) %>% round(4)
CoefOptimp <- if_else(min(coefsall[c(2,4,6),4]) <= 0.05, "*", "")

#Plot comparing the two regresisons


plot3 <- ggplot(traitdf, aes(x=Year, y=Trait,color=Class))+
  ylab(trait)+
  geom_point(size=2)+
  scale_color_brewer(palette = "Dark2")+
  geom_smooth(method="lm",se=FALSE,lwd=2)
print(plot3)

print(residplot)



}



#write.csv(spatsanalysisdf,here(datadir,"GOTH_No_Spats_Analysis.csv"))


```

# ANalysis of within year-loc stdev of yield with stdev scaled on yield magnitude

```{r}
df <- fullweatherentrydf[,c(1:3,7,83)]

genoyrvalues <- read_excel_allsheets(here("Data","Bayer2021ACRERsheet.xlsx"))
genoyrvalues <- genoyrvalues[[3]]
genoyrvalues <- genoyrvalues %>% select(Genotype,Year) %>% mutate(Year = Year - 1982) %>% unique()

df <-  left_join(df,genoyrvalues,by = "Genotype")

df <- df %>% mutate(Class = if_else(total_deficit < 13, "Optimal", if_else(total_deficit < 15, "Mid", "High")))

temp <- df %>% count(Class)
hist(df$total_deficit)

df <- df %>% na.omit() %>% group_by(Genotype, Year, Class) %>% summarize(Yld_Stdev = var(YLD_BE), Yld_Mean = mean(YLD_BE)) %>% 
  na.omit() %>% ungroup() %>% mutate(Yld_varstd = Yld_Stdev/Yld_Mean)



traits = colnames(df[,c(4:6)])

for (t in 3){

trait <- traits[t]

cat("\n\n\\pagebreak\n") 
  cat(paste(trait," With GS as Factor no SPATs"))
  cat("\n")

traitdf <- df %>%  select(c("Class","Genotype","Year",trait)) %>% na.omit() %>% rename("Trait" = trait) %>%
  mutate(Year = as.numeric(Year)) %>% mutate(Class = as.factor(Class))

#Visualize the Data

bxp2 <- ggboxplot(traitdf, x="Class", y="Trait",color = "Class",lwd = 1)+
  ylab(trait)
print(bxp2)

# Check Normality Assumption

normmodel <- lm(Trait~Year+factor(Class)+Year*factor(Class),data = traitdf)
normplot <-  ggqqplot(residuals(normmodel))+
   ggtitle(paste0(trait," QQ Plot"))

print(normplot)

shapiropval <- shapiro.test(normmodel$residuals)
shapiropval <- shapiropval$p.value %>% round(4)


#Constant variance


Homoscedasticity <- ncvTest(normmodel)
Homoscedasticityp <- Homoscedasticity$p %>% round(4)

#If p-value is not significant, variance is constant

#Anova test for variety and replication

anovatest <- anova_test(data = traitdf, Trait~Year+factor(Class)+Year*factor(Class))

anovatest <- as.tibble(anovatest)

# Determine if any of the effects are significant
anovanames <- anovatest %>% pull(Effect)
anovapvals <- anovatest %>% pull(p)

#Fit regression across treatments
modelall <- lm(Trait~ Year + factor(Class)+Year*factor(Class),traitdf)
summaryall <- summary(modelall)

predictions <- predict(modelall,traitdf)
residdf <- data.frame("residuals" = summaryall$residuals, "predictions" = predictions, "year" = traitdf$Year)
residplot <- ggplot(residdf, aes(x = predictions, y=residuals))+
  geom_point()

coefsall <- summaryall$coefficients

CoefHigh <- sum(coefsall[c(2),1]) %>% round(4)
CoefHighp <- if_else(min(coefsall[c(2),4]) <= 0.05, "*", "")

CoefMid <- sum(coefsall[c(2,3,5),1]) %>% round(4)
CoefMidp <- if_else(min(coefsall[c(2,3,5),4]) <= 0.05, "*", "")

CoefOptim <- sum(coefsall[c(2,4,6),1]) %>% round(4)
CoefOptimp <- if_else(min(coefsall[c(2,4,6),4]) <= 0.05, "*", "")

#Plot comparing the two regresisons


plot3 <- ggplot(traitdf, aes(x=Year, y=Trait,color=Class))+
  ylab(trait)+
  geom_point(size=2)+
  scale_color_brewer(palette = "Dark2")+
  geom_smooth(method="lm",se=FALSE,lwd=2)
print(plot3)

print(residplot)



}

#colnames(spatsanalysisdf) <- c("Statistic",traits)

#write.csv(spatsanalysisdf,here(datadir,"GOTH_No_Spats_Analysis.csv"))


```


# ANalysis of Yield Data

```{r}
#write.csv(fullweatherentrydf,here("Data","10GS_Binned_Weather_Data.csv"))

df <- fullweatherentrydf[,c(1:3,7,83)]

genoyrvalues <- read_excel_allsheets(here("Data","Bayer2021ACRERsheet.xlsx"))
genoyrvalues <- genoyrvalues[[3]]
genoyrvalues <- genoyrvalues %>% select(Genotype,Year) %>% mutate(Year = Year - 1982) %>% unique()

df <-  left_join(df,genoyrvalues,by = "Genotype")

df <- df %>% mutate(Class = if_else(total_deficit < 14, "Optimal", if_else(total_deficit < 17, "Mid", "High")))


traits = colnames(df[,c(4:5)])

for (t in 1){

trait <- traits[t]

cat("\n\n\\pagebreak\n") 
  cat(paste(trait," With GS as Factor no SPATs"))
  cat("\n")

traitdf <- df %>%  select(c("Class","Genotype","Year",trait)) %>% na.omit() %>% rename("Trait" = trait) %>%
  mutate(Year = as.numeric(Year)) %>% mutate(Class = as.factor(Class))

#Visualize the Data

bxp2 <- ggboxplot(traitdf, x="Class", y="Trait",color = "Class",lwd = 1)+
  ylab(trait)
print(bxp2)

# Check Normality Assumption

normmodel <- lm(Trait~Year+factor(Class)+Year*factor(Class),data = traitdf)
normplot <-  ggqqplot(residuals(normmodel))+
   ggtitle(paste0(trait," QQ Plot"))

print(normplot)

shapiropval <- shapiro.test(normmodel$residuals)
shapiropval <- shapiropval$p.value %>% round(4)


#Constant variance


Homoscedasticity <- ncvTest(normmodel)
Homoscedasticityp <- Homoscedasticity$p %>% round(4)

#If p-value is not significant, variance is constant

#Anova test for variety and replication

anovatest <- traitdf %>% anova_test(Trait~Year+factor(Class)+Year*factor(Class))

anovatest <- as.tibble(anovatest)

# Determine if any of the effects are significant
anovanames <- anovatest %>% pull(Effect)
anovapvals <- anovatest %>% pull(p)

#Fit regression across treatments
modelall <- lm(Trait~ Year + factor(Class)+Year*factor(Class),traitdf)
summaryall <- summary(modelall)

predictions <- predict(modelall,traitdf)
residdf <- data.frame("residuals" = summaryall$residuals, "predictions" = predictions, "year" = traitdf$Year)
residplot <- ggplot(residdf, aes(x = predictions, y=residuals))+
  geom_point()

coefsall <- summaryall$coefficients

CoefHigh <- sum(coefsall[c(2),1]) %>% round(4)
CoefHighp <- if_else(min(coefsall[c(2),4]) <= 0.05, "*", "")

CoefMid <- sum(coefsall[c(2,3,5),1]) %>% round(4)
CoefMidp <- if_else(min(coefsall[c(2,3,5),4]) <= 0.05, "*", "")

CoefOptim <- sum(coefsall[c(2,4,6),1]) %>% round(4)
CoefOptimp <- if_else(min(coefsall[c(2,4,6),4]) <= 0.05, "*", "")

#Plot comparing the two regresisons


plot3 <- ggplot(traitdf, aes(x=Year, y=Trait,color=Class))+
  ylab(trait)+
  geom_point(size=2)+
  scale_color_brewer(palette = "Dark2")+
  geom_smooth(method="lm",se=FALSE,lwd=2)
print(plot3)

print(residplot)


table <- data.frame("Assumption Checking" = c("Homoscedasticity", "Shapiro-Wilk",rep("",times=2)),
                    "p-values" = c(Homoscedasticityp,shapiropval,rep("",times=2)) ,
                    "." = c(rep("",times=4)),
                    "ANOVA Effects" = c(anovanames),
                    "p-values" = c(anovapvals),
                    ".." = c(rep("",times=4)),
                    "Trt" = c("High","Mid","Optimal",""),
                    "Slope" = c(CoefHigh,CoefMid,CoefOptim,""),
                    "Sig" = c(CoefHighp,CoefMidp,CoefOptimp,""))


print(knitr::kable(table, caption = paste0(trait, " Statistics")))

spatsanalysisdf <- cbind(spatsanalysisdf,c(Homoscedasticityp,shapiropval,anovapvals,CoefTrt0,CoefTrt40,CoefTrt100))

}

colnames(spatsanalysisdf) <- c("Statistic",traits)

#write.csv(spatsanalysisdf,here(datadir,"GOTH_No_Spats_Analysis.csv"))


```
